---
title: "FluVacs Figures - Draft"
author: "Kari and JT"
date: "May 2, 2016"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r,include=F}
require(knitr)
require(plyr)
require(ggplot2)
require(reshape2)
require(ggplot2)
require(ggdendro)
require(grid)
opts_chunk$set(fig.align="center",warning=FALSE,tidy=T, fig.width=5,fig.height=4,cache = T)
theme_set(new = theme_classic()+ theme(
axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))) # to make nice plots
source("../scripts/useful_functions.R") # useful functions adapted largley from HIVE work
```
```{r, variant_calls}

var.2007.8<-read.csv("../data/processed/Run_1293/Variants/all.sum.csv",stringsAsFactors = F)
x<-read.csv("../data/processed/Run_1304/Variants/all.sum.csv",stringsAsFactors = F) # the rest of these samples
var.2007.8<-rbind(var.2007.8,x) # combine both runs
other.seasons<-read.csv("../data/processed/Run_1412/Variants/all.sum.csv",stringsAsFactors = F)


var.2004.5.df<-processing(data.df = other.seasons,meta.df = titer.2004.5,pval = 0.01,phred = 35,mapq = 30,read_cut = c(32,94))

var.2004.5.df<-subset(var.2004.5.df,season=="04-05")
var.2005.6.df<-processing(data.df = other.seasons,meta.df = titer.2005.6,pval = 0.01,phred = 35,mapq = 30,read_cut = c(32,94))
var.2005.6.df<-subset(var.2005.6.df,season=="05-06")

var.2007.8.df<-processing(data.df = var.2007.8,meta.df = titer.2007.8,pval = 0.01,phred = 35,mapq = 30,read_cut = c(32,94))

all.df<-rbind(var.2004.5.df,var.2005.6.df)
all.df<-rbind(all.df,var.2007.8.df)

all.df<-subset(all.df,freq.var>=0.01)
##### Now for the duplicate runs #######

var.2004.5.df2<-read.csv("../data/processed/2004_2005/Variants/all.sum.csv",stringsAsFactors = F)
var.2004.5.df2<-processing(data.df = var.2004.5.df2,meta.df = titer.2004.5,pval = 0.01,phred = 35,mapq = 30,read_cut = c(32,94))

var.2005.6.df2<-read.csv("../data/processed/2005-2006/Variants/all.sum.csv",stringsAsFactors = F)
var.2005.6.df2<-processing(data.df = var.2005.6.df2,meta.df = titer.2005.6,pval = 0.01,phred = 35,mapq = 30,read_cut = c(32,94))

var.2007.8.df2<-read.csv("../data/processed/2007-2008/Variants/all.sum.csv",stringsAsFactors = F)
var.2007.8.df2<-processing(data.df = var.2007.8.df2,meta.df = titer.2007.8,pval = 0.01,phred = 35,mapq = 30,read_cut = c(32,94))

### Join duplicates #####

dups.2004.5<-join_dups(data1.df = var.2004.5.df,data2.df = var.2004.5.df2)
dups.2005.6<-join_dups(data1.df = var.2005.6.df,data2.df = var.2005.6.df2)
dups.2007.8<-join_dups(data1.df = var.2007.8.df,data2.df = var.2007.8.df2)

##### Merge duplicates with intial data that was >1e5

qual.2004.5<-high_qual(data1.df = var.2004.5.df,dups.df = dups.2004.5,titer = 1e3) # only duplicates above 1e3 kept
qual.2005.6<-high_qual(data1.df = var.2005.6.df,dups.df = dups.2005.6,titer = 1e3)
qual.2007.8<-high_qual(data1.df = var.2007.8.df,dups.df = dups.2007.8,titer = 1e3)
all.qual.df<-rbind(qual.2004.5,qual.2005.6)
all.qual.df<-rbind(all.qual.df,qual.2007.8)

all.qual.df<-subset(all.qual.df,freq.var>=0.01)

```


##Figure 3 Coverage plots

These are just from the first runs. I'm not including the duplicates here - should I?

```{r}
cov.2007.8<-cbind(read.csv("../data/processed/Run_1293/deepSNV/all.coverage.csv",stringsAsFactors = F),read.csv("../data/processed/Run_1304/deepSNV/all.coverage.csv",stringsAsFactors = F))

other.seasons.cov<-read.csv("../data/processed/Run_1412/deepSNV/all.coverage.csv",stringsAsFactors = F)


cov.2004.5<-subset(other.seasons.cov,Id %in% titer.2004.5$Sample)
cov.2005.6<-subset(other.seasons.cov,Id %in% titer.2005.6$Sample)
```






# Initial data processing

```{r,titers}
# read in the csvs and add a season column to use later
titer.2004.5<-read.csv("../Titers_status_2004-2005.csv",stringsAsFactors = F)
titer.2004.5$season<-"04-05"
titer.2005.6<-read.csv("../Titers_status_2005-2006.csv",stringsAsFactors = F)
titer.2005.6$season<-"05-06"
titer.2007.8<-read.csv("../Titers_status_2007-2008.csv",stringsAsFactors = F)
titer.2007.8$season<-"07-08"
```



So after all that work we are left with all the samples (all seasons) in all.df and only those variants from high quality data in all.qual.df. This includes samples with >10^5^ genomes per microliter or > 10^3^ genomes per microliter but were sequenced in duplicate.


# Frequency of Variants

```{r}
ggplot(subset(all.df,freq.var<=0.99),aes(x=freq.var))+geom_histogram(color="white")+scale_x_log10() + ggtitle("All samples")
ggplot(subset(all.qual.df,freq.var<=0.9),aes(x=freq.var))+geom_histogram(color="white")+scale_x_log10() + ggtitle("Only high quality samples")

```

# Diversity measures

## SNV count per sample

```{r,snv_sample}
count_muts<-function(data.df){
  ddply(data.df,~Lauring_Id+Vax+season,function(x) dim(x)[1])
}

all.snv<-count_muts(all.df)
qual.snv<-count_muts(all.qual.df)
```

### Whole genome
#### all data

```{r}
ggplot(data=subset(all.snv,!is.na(Vax)),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Number of SNV")
```

#### High quality

```{r}
ggplot(data=subset(qual.snv,!is.na(Vax)),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y',dotsize = 0.7)+ylab("Number of SNV")
```

### HA NA

#### all data

```{r}
all.snv<-count_muts(subset(all.df,chr %in%c("HA","N_A","NR")))
ggplot(data=subset(all.snv,!is.na(Vax)),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Number of SNV")
```

#### High quality

```{r}
qual.snv<-count_muts(subset(all.qual.df,chr %in%c("HA","N_A","NR")))

ggplot(data=subset(qual.snv,!is.na(Vax)),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Number of SNV")
```

### Anitigenic sites

```{r}
Ha_antigenic<-function(df){
  #non_coding<-29
    
  df<-mutate(df,AA_pos= (pos-1)%/%3+1)

  
  df$antigenic<-F
  antigenic_sites<-c(122,124,126,130,131,132,133,135,137,138,140,142,143,144,145,146,150,152,168,128,129,155,156,157,158,159,160,163,164,165,186,187,188,189,190,192,193,194,196,197,198,44,45,46,47,48,50,51,53,54,273,275,276,278,279,280,294,297,299,300,304,305,307,308,309,310,311,312,96,102,103,117,121,167,170,171,172,173,174,175,176,177,179,182,201,203,207,208,209,212,213,214,215,216,217,218,219,226,227,228,229,230,238,240,242,246,247,248,57,59,62,63,67,75,78,80,81,82,83,86,87,88,91,92,94,109,260,261,262,265)
  AA_anti<-antigenic_sites+17
  # antigenic sites:
# Site A: 122,124,126,130,131,132,133,135,137,138,140,142,143,144,145,146,150,152,168
# Site B: 128,129,155,156,157,158,159,160,163,164,165,186,187,188,189,190,192,193,194,196,197,198
# Site C: 44,45,46,47,48,50,51,53,54,273,275,276,278,279,280,294,297,299,300,304,305,307,308,309,310,311,312
# Site D: 96,102,103,117,121,167,170,171,172,173,174,175,176,177,179,182,201,203,207,208,209,212,213,214,215,216,217,218,219,226,227,228,229,230,238,240,242,246,247,248
# Site E: 57,59,62,63,67,75,78,80,81,82,83,86,87,88,91,92,94,109,260,261,262,265
# All sites: 122,124,126,130,131,132,133,135,137,138,140,142,143,144,145,146,150,152,168,128,129,155,156,157,158,159,160,163,164,165,186,187,188,189,190,192,193,194,196,197,198,44,45,46,47,48,50,51,53,54,273,275,276,278,279,280,294,297,299,300,304,305,307,308,309,310,311,312,96,102,103,117,121,167,170,171,172,173,174,175,176,177,179,182,201,203,207,208,209,212,213,214,215,216,217,218,219,226,227,228,229,230,238,240,242,246,247,248,57,59,62,63,67,75,78,80,81,82,83,86,87,88,91,92,94,109,260,261,262,265
# Cluster Transitions: 145,155,156,158,159,189,193
    
    
    
  df$antigenic[df$chr=="HA" & df$AA_pos %in% AA_anti]<-T
  return(df)
}


all.df<-Ha_antigenic(all.df)

all.df.counts<-count_muts(subset(all.df,antigenic==T))

ggplot(data=subset(all.df.counts,!is.na(Vax)),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y',dotsize = 0.8)+ylab("Number of SNV")+ggtitle("all samples")

all.qual.df<-Ha_antigenic(all.qual.df)

all.qual.df.counts<-count_muts(subset(all.qual.df,antigenic==T))
ggplot(data=subset(all.qual.df.counts,!is.na(Vax)),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y',dotsize = 0.8)+ylab("Number of SNV")+ggtitle("quality samples")

all.df.counts<-count_muts(subset(all.df,antigenic==T & freq.var<0.9))

ggplot(data=subset(all.df.counts),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Number of SNV")+ggtitle("all samples frequency<90%")



all.qual.df.counts<-count_muts(subset(all.qual.df,antigenic==T & freq.var<0.9))
ggplot(data=subset(all.qual.df.counts),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Number of SNV")+ggtitle("quality samples (frequency<90%)")


```

## Shannon's Entropy

For the entropy calculations I am assuming 39000 potential variants. This is not right but probably close enough. It is just used to normalize the data. In the future we can use the fasta reference files to get the exact number.

### Whole genome

```{r}
possible_vars<-3900 # fix to be accurate

H<-function(x){
  x<-subset(x,freq.var<0.9)
  H_pos<-ddply(x,~chr+pos,summarize,wt=1-sum(freq.var),H=-sum(c(freq.var*log(freq.var),wt*log(wt))))
sum(H_pos$H)/(possible_vars/3)
}

all.H<-ddply(all.df,~Lauring_Id+Vax+season,H)

ggplot(data=subset(all.H),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Shannon's Entropy")+scale_y_log10()+ggtitle("all samples")


qual.H<-ddply(all.qual.df,~Lauring_Id+Vax+season,H)
ggplot(data=subset(qual.H),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Shannon's Entropy")+scale_y_log10()+ggtitle("quality samples")


```

### HA NA

```{r}
all.HA.NA.H<-ddply(subset(all.df,chr %in% c("HA","N_A","NR")),~Lauring_Id+Vax+season,H)

ggplot(data=subset(all.HA.NA.H),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Shannon's Entropy")+scale_y_log10()+ggtitle("all samples")


qual.HA.NA.H<-ddply(subset(all.qual.df, chr %in% c("HA","N_A","NR")),~Lauring_Id+Vax+season,H)
ggplot(data=subset(qual.H),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Shannon's Entropy")+scale_y_log10()+ggtitle("quality samples")

```

### Anitigenic sites


```{r}

all.ant.H<-ddply(subset(all.df,antigenic==T),~Lauring_Id+Vax+season,H)

ggplot(data=subset(all.ant.H),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Shannon's Entropy")+scale_y_log10()+ggtitle("all samples")




qual.ant.H<-ddply(subset(all.qual.df,antigenic==T),~Lauring_Id+Vax+season,H)

ggplot(data=subset(qual.ant.H),aes(y=V1,x=Vax))+geom_dotplot(stackdir = "center",binaxis = 'y')+ylab("Shannon's Entropy")+scale_y_log10()+ggtitle("quality samples")



```



#Heat map
Theses are made with the quality samples

```{r,fig.height=10,fig.width=7}
make_heat_map(subset(all.qual.df,season=="07-08"))

make_heat_map(subset(all.qual.df,season=="05-06"))
make_heat_map(subset(all.qual.df,season=="04-05"))

```

# Antigenic Sites

### All data
```{r,}
#kable(subset(all.df,antigenic==T & season=="07-08", select=c(Lauring_Id,mutation,freq.var,AA_Pos)))

```

### High quality
```{r}
x<-Ha_antigenic(qual.2007.8)

kable(subset(x,antigenic==T& season=="07-08", select=c(Lauring_Id,mutation,freq.var,AA_pos)))

```