---
title: "FluVacs Figures"
author: "JT"
date: "December 29, 2016"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    
---

```{r,include=F}
require(knitr)
require(plyr)
require(ggplot2)
require(reshape2)
require(ggplot2)
require(ggdendro)
require(grid)
opts_chunk$set(fig.align="center",warning=FALSE,tidy=T,cache = F,echo=F)
theme_set(new = theme_classic()+ theme(
axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))) # to make nice plots
source("../scripts/useful_functions.R") # useful functions adapted largley from HIVE work
```

```{r}
minor.07.08<-read.csv("../results/2007-2008.wg.csv")
```

Here verify the average coverage of the sample is >1000

```{r}
read_rbind<-function(list,list2=NULL){
  out<-data.frame()
  for(i in 1:length(list)){
    print(paste0("reading in ",list[i]))
    
    x<-read.csv(list[i],stringsAsFactors = F)
    if(is.null(list2)==F){
      x$run=list2[i]
      print(paste0("appending run column: ",list2[i]))
    }
    out<-rbind(x,out)
  }
  return(out)
}


cov_files<-c("../data/processed/Run_1293/deepSNV/all.coverage.csv","../data/processed/Run_1304/deepSNV/all.coverage.csv","../data/processed/Run_1412/deepSNV/all.coverage.csv","../data/processed/2004-2005/deepSNV/all.coverage.csv","../data/processed/2005-2006/deepSNV/all.coverage.csv","../data/processed/2007-2008/deepSNV/all.coverage.csv")


 cov<-read_rbind(cov_files,c("1293","1304","1412","2004-5","2005-6","2007-8"))
 
 cov.sample<-ddply(cov,~Id+run,summarize,cov=mean(coverage))

 ggplot(cov.sample,aes(x=as.factor(run),y=cov))+geom_boxplot()+geom_point(position="jitter")+xlab("Run")+ylab("Mean Coverage in the sample")+scale_y_log10() #+ scale_y_continuous(limits=c(0,50000),breaks = seq(0,50000,by=10000))

 poor_cov<-subset(cov.sample,cov<1000)
 poor_cov$Id<-as.numeric(poor_cov$Id)
removed<-dim(subset(minor.07.08,Id %in% poor_cov$Id))[1]
 
```
In the 2007-2008 data set there were `r removed` samples removed from the the minority analysis based on a mean coverage less than 1000X.

# Figure 3

High quality is >10^5^ or >10^3^ sequenced in duplicate.
Currently I am making these plots with just the 2007-2008 samples. 

These are the variants between 1-50%. each bin is 1% wide. 
```{r}
#ggplot(subset(var.2007.8.df,freq.var<=0.5),aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)#+scale_y_log10() + ggtitle("All samples")

fig_3A<-ggplot(minor.07.08,aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)+xlab("Frequency")+ylab("Number of iSNV")+scale_x_continuous(breaks=seq(0,.5,0.1),labels=c("0%","10%","20%","30%","40%","50%"))
#fig_3A+ggtitle("Figure 3A")

pdf("./figures/fig_3A.pdf")
fig_3A
dev.off()


minor_counts<-ddply(minor.07.08,~mutation,summarize,count=length(unique(Id)))

fig_3B<-ggplot(minor_counts,aes(x=count))+geom_histogram(color="white",binwidth=1)+xlab("Samples with iSNV")+ylab("Number of iSNV")
#fig_3B+ggtitle("Figure 3B")



pdf("./figures/fig_3B.pdf")
fig_3B
dev.off()


require(cowplot)
```

The x axis is so large because there are 2 mutations that are found in 62 and 41 samples. Both of these are infered minor variants.


```{r}
# XAI.cut is T if the titer is above 40
qual.ha.snv<-ddply(minor.07.08,~HAI.cut+Id,summarize,HAI=ifelse(unique(HAI.cut)==F,yes = "< 40",no= "\u2265 40"),iSNV=length(which(chr=="HA")))
# If there are no snv in HA then length returns 0. HAI.cut is true if the titer is >=40

## define custom median function https://designdatadecisions.wordpress.com/2015/06/09/graphs-in-r-overlaying-data-summaries-in-dotplots/
plot.median <- function(x) {
  m <- median(x)
  c(y = m, ymin = m, ymax = m)
}


fig_3C<-ggplot(data=qual.ha.snv,aes(y=iSNV,x=HAI))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.09)+ylab("Number of iSNV")+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.95, size=0.5)+xlab("HAI titer")
#fig_3C
write.csv(x = qual.ha.snv,"./2007-2008.HA.HAI.stats.csv")
pdf("./figures/fig_3C.pdf")
fig_3C
dev.off()

qual.na.snv<-ddply(minor.07.08,~NAI.cut+Id,summarize,NAI=ifelse(unique(NAI.cut)==F,yes = "< 40",no= "\u2265 40"),iSNV=length(which(chr =="NR")))
qual.na.snv<-subset(qual.na.snv,!(is.na(NAI)))

fig_3D<-ggplot(data=qual.na.snv,aes(y=iSNV,x=NAI))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.05)+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.95, size=0.5)+ylab("Number of iSNV")+xlab("NAI titer")
#fig_3D
write.csv(x = qual.na.snv,"./2007-2008.NA.NAI.stats.csv")

pdf("./figures/fig_3D.pdf")
fig_3D
dev.off()
```
```{r,fig_3,fig.width=10,fig.height=10}
require(cowplot)
plot_grid(fig_3A, fig_3B, fig_3C, fig_3D, labels = c("A", "B", "C", "D"), ncol = 2)

```


There are very few iSNV in NA. 


# Table 3
Average iSNV/ segment

All data 
```{r}
# The problem here is that it does not include the cases where no variants are found.
# These add_missing functions correct that.
add_missing.intervention<-function(x){ #This is a helper function to add the cases where there were no variants called.
missing_chr<-levels(x$chr)[!((levels(x$chr) %in% as.character(x$chr)))]
  if(length(missing_chr)>0){
    for(i in 1:length(missing_chr)){
      x<-rbind(x,data.frame(Intervention=unique(x$Intervention),chr=missing_chr[i],Id=unique(x$Id),iSNV=0,n=unique(x$n)))
    }
  }
  return(x)
}

add_missing.HAI<-function(x){ #This is a helper function to add the cases where there were no variants called.
  missing_chr<-levels(x$chr)[!((levels(x$chr) %in% as.character(x$chr)))]
  if(length(missing_chr)>0){
    for(i in 1:length(missing_chr)){
      x<-rbind(x,data.frame(HAI.cut=unique(x$HAI.cut),chr=missing_chr[i],Id=unique(x$Id),iSNV=0,n=unique(x$n)))
    }
  }  
  return(x)
}

count_tables<-function(data,return=F){
intervention.chr.counts<-ddply(data,~Intervention+chr+Id,summarize,iSNV=length(mutation))

intervention.chr.counts$chr<-factor(intervention.chr.counts$chr,levels=c("PB2","PB1","PA","HA","NP","NR","M","NS"))


columns<-unique(intervention.chr.counts$Intervention)
intervention.chr.counts$n=0
for(i in 1:length(columns)){ # How many samples are there in each intervention
x<-length(unique(intervention.chr.counts$Id[which(intervention.chr.counts$Intervention==columns[i])]))
intervention.chr.counts$n[which(intervention.chr.counts$Intervention==columns[i])]=x
}

intervention.chr.counts<-mutate(intervention.chr.counts,Intervention=paste0(Intervention," ","(n=",n,")"))

# I need to add the cases where no variants were found.

intervention.chr.counts<-ddply(intervention.chr.counts,~Id,add_missing.intervention)

intervention.chr.sum<-ddply(intervention.chr.counts,~Intervention+chr,summarize,mean_iSNV=mean(iSNV),range=IQR(iSNV))

intervention.chr.tab<-mutate(intervention.chr.sum,iSNV=paste0(round(mean_iSNV,2)," \u00b1 ","(",range,")"))
intervention.chr.table<-dcast(intervention.chr.tab,chr~Intervention,value.var="iSNV")
#knitr::kable(qual.chr.table)

HAI.chr.counts<-ddply(data,~HAI.cut+chr+Id,summarize,iSNV=length(mutation))
HAI.chr.counts$chr<-factor(HAI.chr.counts$chr,levels=c("PB2","PB1","PA","HA","NP","NR","M","NS"))

HAI.chr.counts<-mutate(HAI.chr.counts,HAI.cut=ifelse(HAI.cut,yes = "\u2265 40",no= "< 40"))

columns<-unique(HAI.chr.counts$HAI.cut)
HAI.chr.counts$n=0
for(i in 1:length(columns)){
x<-length(unique(HAI.chr.counts$Id[which(HAI.chr.counts$HAI.cut==columns[i])]))
HAI.chr.counts$n[which(HAI.chr.counts$HAI.cut==columns[i])]=x
}

HAI.chr.counts<-mutate(HAI.chr.counts,HAI.cut=paste0(HAI.cut," ","(n=",n,")"))

HAI.chr.counts<-ddply(HAI.chr.counts,~Id,add_missing.HAI)

HAI.chr.sum<-ddply(HAI.chr.counts,~HAI.cut+chr,summarize,mean_iSNV=mean(iSNV),range=IQR(iSNV))

HAI.chr.tab<-mutate(HAI.chr.sum,iSNV=paste0(round(mean_iSNV,2)," \u00b1 ","(",range,")"))
HAI.chr.table<-dcast(HAI.chr.tab,chr~HAI.cut,value.var="iSNV")
#knitr::kable(HAI.chr.table)  
all<-join_all(dfs = list(HAI.chr.table,intervention.chr.table),by="chr")

all$chr<-factor(all$chr,levels = c("PB2","PB1","PA","HA","NP","NR","M","NS"))

all<-all[order(all$chr,decreasing = F),]

if(return==T){
  return(list(HAI.chr.counts,intervention.chr.counts))
}
knitr::kable(all)


}

count_tables(minor.07.08)
```
```{r}
write.csv("./table2_data.csv",x = inter)
```

# Figure 4 

# Figure 4 

```{r, isnv_by_day}
isnv.day.intervention<-ddply(minor.07.08,~Intervention+Id+Day.of.Infection.sample.collected,summarize,iSNV=length(mutation))


fig_4A<-ggplot(data=isnv.day.intervention,aes(y=iSNV,x=factor(Day.of.Infection.sample.collected),fill=factor(Intervention)))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.4,position='dodge')+geom_boxplot(aes(lower=..middle.., upper=..middle..,ymin=..middle.., ymax=..middle..),outlier.shape = 32)+ylab("Number of iSNV")+xlab("Day of Infection")+scale_fill_discrete(name="Intervention")
#fig_4A

isnv.day.HAI<-ddply(minor.07.08,~HAI.cut+Id+Day.of.Infection.sample.collected,summarize,HAI=ifelse(sum(HAI.cut)==0,yes = "< 40",no= "\u2265 40"),iSNV=length(mutation))


fig_4B<-ggplot(data=isnv.day.HAI,aes(y=iSNV,x=factor(Day.of.Infection.sample.collected),fill=factor(HAI)))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.4,position='dodge')+geom_boxplot(aes(lower=..middle.., upper=..middle..,ymin=..middle.., ymax=..middle..),outlier.shape = 32)+ylab("Number of iSNV")+xlab("Day of Infection")+scale_fill_discrete(name="HAI")
#fig_4B

isnv.day.NAI<-ddply(minor.07.08,~NAI.cut+Id+Day.of.Infection.sample.collected,summarize,NAI=ifelse(sum(NAI.cut)==0,yes = "< 40",no= "\u2265 40"),iSNV=length(mutation))

isnv.day.NAI<-subset(isnv.day.NAI,!(is.na(NAI)))

fig_4C<-ggplot(data=isnv.day.NAI,aes(y=iSNV,x=factor(Day.of.Infection.sample.collected),fill=factor(NAI)))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.4,position='dodge')+geom_boxplot(aes(lower=..middle.., upper=..middle..,ymin=..middle.., ymax=..middle..),outlier.shape = 32)+ylab("Number of iSNV")+xlab("Day of Infection")+scale_fill_discrete(name="NAI")
#fig_4C

#stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,geom = "crossbar", width = 0.5)


pdf("./figures/fig_4A.pdf",width = 20,height = 12)
fig_4A
dev.off()

pdf("./figures/fig_4B.pdf",width = 20,height = 12)
fig_4B
dev.off()

pdf("./figures/fig_4C.pdf",width = 20,height = 12)
fig_4C
dev.off()

```


```{r,fig_4,fig.height=10}
plot_grid(fig_4A, fig_4B, fig_4C, labels = c("A", "B", "C"), ncol = 1)

```

#Supplemental Figure 3)

These are just from the first runs. These have a sliding window of 100 with a step of 100 no overlap.

```{r,coverage}

titer.2004.5<-read.csv("../Titers_status_2004-2005.csv",stringsAsFactors = F)
titer.2004.5$season<-"04-05"
titer.2005.6<-read.csv("../Titers_status_2005-2006.csv",stringsAsFactors = F)
titer.2005.6$season<-"05-06"
titer.2007.8<-read.csv("../Titers_status_2007-2008.csv",stringsAsFactors = F)
titer.2007.8$season<-"07-08"

cov.2007.8<-rbind(read.csv("../data/processed/Run_1293/deepSNV/all.coverage.csv",stringsAsFactors = F),read.csv("../data/processed/Run_1304/deepSNV/all.coverage.csv",stringsAsFactors = F))

other.seasons.cov<-read.csv("../data/processed/Run_1412/deepSNV/all.coverage.csv",stringsAsFactors = F)


cov.2004.5<-subset(other.seasons.cov,Id %in% titer.2004.5$Sample)
cov.2005.6<-subset(other.seasons.cov,Id %in% titer.2005.6$Sample)


cov.plot.04.05<-cov_plot(cov.2004.5,"2004-2005",100,100)
cov.plot.04.05

cov.plot.05.06<-cov_plot(cov.2005.6,"2005-2006",100,100)
cov.plot.05.06

cov.plot.07.08<-cov_plot(cov.2007.8,"2007-2008",100,100)
cov.plot.07.08

pdf("./figures/sup_1A.pdf")
cov.plot.07.08
dev.off()

pdf("./figures/sup_1B.pdf")
cov.plot.04.05
dev.off()

pdf("./figures/sup_1C.pdf")
cov.plot.05.06
dev.off()
```

If we plot on a log scale the bars are well above 0. 


So there are a few samples with poor coverage that should be removed from the analysis, but none of them were in the 2007-2008 year.


```{r}
concat.pos<-ddply(cov.2007.8,~chr, summarize, low=min(concat.pos), high=max(concat.pos))

```

# Supplemental 5

```{r}
meta.all<-read.csv("../data/raw/all.meta.csv",stringsAsFactors = F)
titer_data<-subset(meta.all,select=c(Id,season,HAI.post.vax,NAI.post.vax))
write.csv(x = titer_data,file = "./titer_data.csv")
```


Diversity by titer

```{r}
qual.ha.snv<-ddply(minor.07.08,~HAI.post.vax+Id,summarize,HAI=HAI.post.vax[1],iSNV=length(which(chr=="HA")))

## define custom median function https://designdatadecisions.wordpress.com/2015/06/09/graphs-in-r-overlaying-data-summaries-in-dotplots/
plot.median <- function(x) {
  m <- median(x)
  c(y = m, ymin = m, ymax = m)
}


sup_5A<-ggplot(data=qual.ha.snv,aes(y=iSNV,x=as.factor(HAI)))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.1)+ylab("Number of iSNV")+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.95, size=0.5)+xlab("HAI")+scale_x_discrete(labels=paste0("1:",unique(qual.ha.snv$HAI)))+scale_y_continuous(breaks=c(0,2,4,6))
#titer_3C

pdf("./figures/sup_5A.pdf",width = 10)
sup_5A
dev.off()

qual.na.snv<-ddply(minor.07.08,~NAI.post.vax+Id,summarize,NAI=NAI.post.vax[1],iSNV=length(which(chr =="NR")))
qual.na.snv<-subset(qual.na.snv,!(is.na(NAI)))

sup_5B<-ggplot(data=qual.na.snv,aes(y=iSNV,x=as.factor(NAI)))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.05)+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.95, size=0.5)+ylab("Number of iSNV")+xlab("NAI")+scale_x_discrete(labels=paste0("1:",unique(qual.na.snv$NAI)))+scale_y_continuous(breaks=c(0,2,4,6,8,10))
#titer_3D

pdf("./figures/sup_5B.pdf",width=12)
sup_5B
dev.off()
```

```{r,fig.width=20}
plot_grid(sup_5A, sup_5B, labels = c("A", "B"), ncol = 2)

```

# Supplemental Figure 6)

Genome wide SNV
```{r,extra_iSNV}
qual.ha.snv<-ddply(minor.07.08,~HAI.cut+Id,summarize,HAI=ifelse(unique(HAI.cut)==F,yes = "< 40",no= "\u2265 40"),iSNV=length(mutation))
# Again if there are no mutations the length above would output 0.

sup_6A<-ggplot(data=qual.ha.snv,aes(y=iSNV,x=HAI))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.3)+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.5, size=0.5)+ylab("Number of iSNV")
#sup_6A
write.csv(x = qual.ha.snv,"./2007-2008.wg.HAI.stats.csv")


qual.na.snv<-ddply(minor.07.08,~NAI.cut+Id,summarize,NAI=ifelse(unique(NAI.cut)==F,yes = "< 40",no= "\u2265 40"),iSNV=length(mutation))
qual.na.snv<-subset(qual.na.snv,!(is.na(NAI)))

sup_6B<-ggplot(data=qual.na.snv,aes(y=iSNV,x=NAI))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.3)+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.5, size=0.5)+ylab("Number of iSNV")
#sup_6B
write.csv(x = qual.na.snv,"./2007-2008.wg.NAI.stats.csv")

# By intervention
qual.intervention.snv<-ddply(minor.07.08,~Intervention+Id,summarize,iSNV=length(mutation))

write.csv(x = qual.intervention.snv,"./2007-2008.wg.intervention.stats.csv")


pdf("./figures/sup_6A.pdf")
sup_6A
dev.off()

pdf("./figures/sup_6B.pdf")
sup_6B
dev.off()

```

```{r}
plot_grid(sup_6A, sup_6B, labels = c("A", "B"), ncol = 2)

```

## Supplemental figure 7 2004-2005 & 2005-2006


```{r}
#ggplot(subset(var.2007.8.df,freq.var<=0.5),aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)#+scale_y_log10() + ggtitle("All samples")

minor.04.05<-read.csv("../results/2004-2005.wg.csv",stringsAsFactors = F)
sup_7A<-ggplot(minor.04.05,aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)+xlab("Frequency")+ylab("Number of iSNV")+scale_x_continuous(breaks=seq(0,.5,0.1),labels=c("0%","10%","20%","30%","40%","50%"))
#sup_7A+ggtitle("2004-2005")

pdf("./figures/sup_7A.pdf")
sup_7A
dev.off()


minor_counts<-ddply(minor.04.05,~mutation,summarize,count=length(unique(Id)))

sup_7B<-ggplot(minor_counts,aes(x=count))+geom_histogram(color="white",binwidth=1)+xlab("Samples with iSNV")+ylab("Number of iSNV")
#sup_7B+ggtitle("2004-2005")



pdf("./figures/sup_7B.pdf")
sup_7B
dev.off()


```



```{r}
#ggplot(subset(var.2007.8.df,freq.var<=0.5),aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)#+scale_y_log10() + ggtitle("All samples")

minor.05.06<-read.csv("../results/2005-2006.wg.csv",stringsAsFactors = F)
sup_7C<-ggplot(minor.05.06,aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)+xlab("Frequency")+ylab("Number of iSNV")+scale_x_continuous(breaks=seq(0,.5,0.1),labels=c("0%","10%","20%","30%","40%","50%"))
#sup_7C+ggtitle("2005-2006")

pdf("./figures/sup_7C.pdf")
sup_7C
dev.off()


minor_counts<-ddply(minor.05.06,~mutation,summarize,count=length(unique(Id)))

sup_7D<-ggplot(minor_counts,aes(x=count))+geom_histogram(color="white",binwidth=1)+xlab("Samples with iSNV")+ylab("Number of iSNV")
#sup_7D+ggtitle("2005-2006")



pdf("./figures/sup_7D.pdf")
sup_7D
dev.off()
```

```{r,fig.width=10,fig.height=10}
plot_grid(sup_7A,sup_7B,sup_7C,sup_7D, labels = c("A", "B", "C", "D"), ncol = 2)

```

# Supplemental tables

## 2004-2005

```{r}
count_tables(minor.04.05)
```

## 2005-2006

```{r}
count_tables(minor.05.06)
```


```{r,eval=F}

check_titers<-function(data){
  H<-mutate(data,H=HAI.geo+HAI.cut)
  print("HAI not equal")
  print(unique(H$Id[H$H==1]))
  
  N<-subset(data,N=NAI.geo+NAI.cut)
  print("NAI not equal")
  print(unique(N$Id[N$N==1]))
  
}

check_titers(minor.04.05)
check_titers(minor.05.06)
check_titers(minor.07.08)
```
