---
title: "FluVacs Figures - Draft"
author: "JT"
date: "June 3, 2016"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    
---

```{r,include=F}
require(knitr)
require(plyr)
require(ggplot2)
require(reshape2)
require(ggplot2)
require(ggdendro)
require(grid)
opts_chunk$set(fig.align="center",warning=FALSE,tidy=T,cache = F,echo=F)
theme_set(new = theme_classic()+ theme(
axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))) # to make nice plots
source("../scripts/useful_functions.R") # useful functions adapted largley from HIVE work
```

```{r}
minor.07.08<-read.csv("../results/2007-2008.wg.csv")
```
# Figure 4

High quality is >10^5^ or >10^3^ sequenced in duplicate.
Currently I am making these plots with just the 2007-2008 samples. Once I have the meta datafor the other seasons it will be trivial to add them (if we want to do that) 

## A) Variant Frequencies

These are the variants between 1-50%. each bin is 1% wide. The y axis is log scaled which in ggplot leads to a problem when there are no observations (lines below axis). I can remove these before publication.
```{r}
#ggplot(subset(var.2007.8.df,freq.var<=0.5),aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)#+scale_y_log10() + ggtitle("All samples")

fig_4A<-ggplot(minor.07.08,aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)+xlab("Frequency")+ylab("Number of iSNV")+scale_x_continuous(breaks=seq(0,.5,0.1),labels=c("0%","10%","20%","30%","40%","50%"))
fig_4A+ggtitle("Figure 4A")

pdf("./figures/fig_4A.pdf")
fig_4A
dev.off()


minor_counts<-ddply(minor.07.08,~mutation,summarize,count=length(mutation))

fig_4B<-ggplot(minor_counts,aes(x=count))+geom_histogram(color="white",binwidth=1)+xlab("Samples with iSNV")+ylab("Number of iSNV")
fig_4B+ggtitle("Figure 4B")



pdf("./figures/fig_4B.pdf")
fig_4B
dev.off()
```

The x axis is so large because there are 2 mutations that are found in 62 and 41 samples. Both of these are infered minor variants. They may be artifacts of our analysis.




```{r}
qual.ha.snv<-ddply(minor.07.08,~HAI.geo+Id,summarize,HAI=ifelse(sum(HAI.geo)==0,yes = "Below geometric mean",no= "Above geomtric mean"),iSNV=sum(which(chr=="HA")))

## define custom median function https://designdatadecisions.wordpress.com/2015/06/09/graphs-in-r-overlaying-data-summaries-in-dotplots/
plot.median <- function(x) {
  m <- median(x)
  c(y = m, ymin = m, ymax = m)
}


fig_4C<-ggplot(data=qual.ha.snv,aes(y=iSNV,x=HAI))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.15)+ylab("Number of iSNV")+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.95, size=0.5)
fig_4C

pdf("./figures/fig_4C.pdf")
fig_4C
dev.off()

qual.na.snv<-ddply(minor.07.08,~NAI.geo+Id,summarize,NAI=ifelse(sum(NAI.geo)==0,yes = "Below geometric mean",no= "Above geomtric mean"),iSNV=sum(which(chr =="NR")))
qual.na.snv<-subset(qual.na.snv,!(is.na(NAI)))

fig_4D<-ggplot(data=qual.na.snv,aes(y=iSNV,x=NAI))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.15)+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.95, size=0.5)+ylab("Number of iSNV")
fig_4D

pdf("./figures/fig_4D.pdf")
fig_4D
dev.off()
```


There are very few iSNV in NA. 


# Table 3
Average iSNV/ segment

All data 
```{r}

qual.chr.counts<-ddply(minor.07.08,~Intervention+chr+Id,summarize,iSNV=length(mutation))

columns<-unique(qual.chr.counts$Intervention)
qual.chr.counts$n=NA
for(i in 1:length(columns)){
x<-length(unique(qual.chr.counts$Id[which(qual.chr.counts$Intervention==columns[i])]))
qual.chr.counts$n[which(qual.chr.counts$Intervention==columns[i])]=x
}

qual.chr.counts<-mutate(qual.chr.counts,Intervention=paste0(Intervention," ","(n=",n,")"))

qual.chr.sum<-ddply(qual.chr.counts,~Intervention+chr,summarize,mean_iSNV=mean(iSNV),range=IQR(iSNV))

qual.chr.tab<-mutate(qual.chr.sum,iSNV=paste0(round(mean_iSNV,2)," \u00b1 ","(",range,")"))
qual.chr.table<-dcast(qual.chr.tab,chr~Intervention,value.var="iSNV")
knitr::kable(qual.chr.table)

```


# iSNV/day Figure 


```{r, isnv_by_day}
isnv.day.intervention<-ddply(minor.07.08,~Intervention+Id+Day.of.Infection.sample.collected,summarize,iSNV=length(mutation))


fig_xA<-ggplot(data=isnv.day.intervention,aes(y=iSNV,x=factor(Day.of.Infection.sample.collected),fill=factor(Intervention)))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.4,position='dodge')+geom_boxplot(aes(lower=..middle.., upper=..middle..,ymin=..middle.., ymax=..middle..),outlier.shape = 32)+ylab("Number of iSNV")+xlab("Day of Infection")+scale_fill_discrete(name="Intervention")
fig_xA

isnv.day.HAI<-ddply(minor.07.08,~HAI.geo+Id+Day.of.Infection.sample.collected,summarize,HAI=ifelse(sum(HAI.geo)==0,yes = "Below geometric mean",no= "Above geomtric mean"),iSNV=length(mutation))


fig_xB<-ggplot(data=isnv.day.HAI,aes(y=iSNV,x=factor(Day.of.Infection.sample.collected),fill=factor(HAI)))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.4,position='dodge')+geom_boxplot(aes(lower=..middle.., upper=..middle..,ymin=..middle.., ymax=..middle..),outlier.shape = 32)+ylab("Number of iSNV")+xlab("Day of Infection")+scale_fill_discrete(name="HAI")
fig_xB

isnv.day.NAI<-ddply(minor.07.08,~NAI.geo+Id+Day.of.Infection.sample.collected,summarize,NAI=ifelse(sum(NAI.geo)==0,yes = "Below geometric mean",no= "Above geomtric mean"),iSNV=length(mutation))

subset(isnv.day.NAI,(is.na(NAI)))
isnv.day.NAI<-subset(isnv.day.NAI,!(is.na(NAI)))

fig_xC<-ggplot(data=isnv.day.NAI,aes(y=iSNV,x=factor(Day.of.Infection.sample.collected),fill=factor(NAI)))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.4,position='dodge')+geom_boxplot(aes(lower=..middle.., upper=..middle..,ymin=..middle.., ymax=..middle..),outlier.shape = 32)+ylab("Number of iSNV")+xlab("Day of Infection")+scale_fill_discrete(name="NAI")
fig_xC

#stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,geom = "crossbar", width = 0.5)


pdf("./figures/fig_xA.pdf",width = 20,height = 12)
fig_xA
dev.off()

pdf("./figures/fig_xB.pdf",width = 20,height = 12)
fig_xB
dev.off()

pdf("./figures/fig_xC.pdf",width = 20,height = 12)
fig_xC
dev.off()



```

#Supplemental Figure 1)

These are just from the first runs. I'm not including the duplicates here - should I? These have a sliding window of 100 with a step of 100 no overlap.

```{r,coverage}

titer.2004.5<-read.csv("../Titers_status_2004-2005.csv",stringsAsFactors = F)
titer.2004.5$season<-"04-05"
titer.2005.6<-read.csv("../Titers_status_2005-2006.csv",stringsAsFactors = F)
titer.2005.6$season<-"05-06"
titer.2007.8<-read.csv("../Titers_status_2007-2008.csv",stringsAsFactors = F)
titer.2007.8$season<-"07-08"

cov.2007.8<-rbind(read.csv("../data/processed/Run_1293/deepSNV/all.coverage.csv",stringsAsFactors = F),read.csv("../data/processed/Run_1304/deepSNV/all.coverage.csv",stringsAsFactors = F))

other.seasons.cov<-read.csv("../data/processed/Run_1412/deepSNV/all.coverage.csv",stringsAsFactors = F)


cov.2004.5<-subset(other.seasons.cov,Id %in% titer.2004.5$Sample)
cov.2005.6<-subset(other.seasons.cov,Id %in% titer.2005.6$Sample)


cov.plot.04.05<-cov_plot(cov.2004.5,"2004-2005",100,100)
cov.plot.04.05

cov.plot.05.06<-cov_plot(cov.2005.6,"2005-2006",100,100)
cov.plot.05.06

cov.plot.07.08<-cov_plot(cov.2007.8,"2007-2008",100,100)
cov.plot.07.08

pdf("./figures/sup_1A.pdf")
cov.plot.07.08
dev.off()

pdf("./figures/sup_1B.pdf")
cov.plot.04.05
dev.off()

pdf("./figures/sup_1C.pdf")
cov.plot.05.06
dev.off()
```

If we plot on a log scale the bars are well above 0. 


So there are a few samples with poor coverage that should be removed from the analysis, but none of them were in the 2007-2008 year.

```{r}
cov.plot.07.08+scale_y_log10(breaks=c(1,10,100,1000,10000,100000))
```



```{r}
concat.pos<-ddply(cov.2007.8,~chr, summarize, low=min(concat.pos), high=max(concat.pos))

```

# Supplemental Figure 2)
```{r,extra_iSNV}
qual.ha.snv<-ddply(minor.07.08,~HAI.geo+Id,summarize,HAI=ifelse(sum(HAI.geo)==0,yes = "Below geometric mean",no= "Above geomtric mean"),iSNV=length(mutation))

sup_2A<-ggplot(data=qual.ha.snv,aes(y=iSNV,x=HAI))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.3)+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.5, size=0.5)+ylab("Number of iSNV")
sup_2A


qual.na.snv<-ddply(minor.07.08,~NAI.geo+Id,summarize,NAI=ifelse(sum(NAI.geo)==0,yes = "Below geometric mean",no= "Above geomtric mean"),iSNV=length(mutation))
qual.na.snv<-subset(qual.na.snv,!(is.na(NAI)))

sup_2B<-ggplot(data=qual.na.snv,aes(y=iSNV,x=NAI))+geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 1,dotsize = 0.3)+stat_summary(fun.data="plot.median", geom="errorbar", colour="black", width=0.5, size=0.5)+ylab("Number of iSNV")
sup_2B


pdf("./figures/sup_2A.pdf")
sup_2A
dev.off()

pdf("./figures/sup_2B.pdf")
sup_2B
dev.off()


```

## Supplemental figure 3 2004-2005


```{r}
#ggplot(subset(var.2007.8.df,freq.var<=0.5),aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)#+scale_y_log10() + ggtitle("All samples")

minor.04.05<-read.csv("../results/2004-2005.wg.csv",stringsAsFactors = F)
sup_3A<-ggplot(minor.04.05,aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)+xlab("Frequency")+ylab("Number of iSNV")+scale_x_continuous(breaks=seq(0,.5,0.1),labels=c("0%","10%","20%","30%","40%","50%"))
sup_3A+ggtitle("2004-2005")

pdf("./figures/sup_3A.pdf")
sup_3A
dev.off()


minor_counts<-ddply(minor.04.05,~mutation,summarize,count=length(mutation))

sup_3B<-ggplot(minor_counts,aes(x=count))+geom_histogram(color="white",binwidth=1)+xlab("Samples with iSNV")+ylab("Number of iSNV")
sup_3B+ggtitle("2004-2005")



pdf("./figures/sup_3B.pdf")
sup_3B
dev.off()



qual.chr.counts<-ddply(minor.04.05,~Intervention+chr+Id,summarize,iSNV=length(mutation))

columns<-unique(qual.chr.counts$Intervention)
qual.chr.counts$n=NA
for(i in 1:length(columns)){
x<-length(unique(qual.chr.counts$Id[which(qual.chr.counts$Intervention==columns[i])]))
qual.chr.counts$n[which(qual.chr.counts$Intervention==columns[i])]=x
}

qual.chr.counts<-mutate(qual.chr.counts,Intervention=paste0(Intervention," ","(n=",n,")"))


qual.chr.sum<-ddply(qual.chr.counts,~Intervention+chr,summarize,mean_iSNV=mean(iSNV),range=IQR(iSNV))

qual.chr.tab<-mutate(qual.chr.sum,iSNV=paste0(round(mean_iSNV,2)," \u00b1 ","(",range,")"))
qual.chr.table<-dcast(qual.chr.tab,chr~Intervention,value.var="iSNV")
knitr::kable(qual.chr.table)

qual.chr.counts<-ddply(minor.04.05,~HAI.geo+chr+Id,summarize,iSNV=length(mutation))

qual.chr.counts<-mutate(qual.chr.counts,HAI.geo=ifelse(HAI.geo,yes = "Above geometric mean",no= "Below geometric mean"))

columns<-unique(qual.chr.counts$HAI.geo)
qual.chr.counts$n=NA
for(i in 1:length(columns)){
x<-length(unique(qual.chr.counts$Id[which(qual.chr.counts$HAI.geo==columns[i])]))
qual.chr.counts$n[which(qual.chr.counts$HAI.geo==columns[i])]=x
}

qual.chr.counts<-mutate(qual.chr.counts,HAI.geo=paste0(HAI.geo," ","(n=",n,")"))


qual.chr.sum<-ddply(qual.chr.counts,~HAI.geo+chr,summarize,mean_iSNV=mean(iSNV),range=IQR(iSNV))

qual.chr.tab<-mutate(qual.chr.sum,iSNV=paste0(round(mean_iSNV,2)," \u00b1 ","(",range,")"))
qual.chr.table<-dcast(qual.chr.tab,chr~HAI.geo,value.var="iSNV")
knitr::kable(qual.chr.table)

```

## Supplemental figure 4 2005-2006


```{r}
#ggplot(subset(var.2007.8.df,freq.var<=0.5),aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)#+scale_y_log10() + ggtitle("All samples")

minor.05.06<-read.csv("../results/2005-2006.wg.csv",stringsAsFactors = F)
sup_4A<-ggplot(minor.05.06,aes(x=freq.var))+geom_histogram(color="white",binwidth = 0.01)+xlab("Frequency")+ylab("Number of iSNV")+scale_x_discrete(breaks=seq(0,.5,0.1),labels=c("0%","10%","20%","30%","40%","50%"))
sup_4A+ggtitle("2005-2006")

pdf("./figures/sup_4A.pdf")
sup_4A
dev.off()


minor_counts<-ddply(minor.05.06,~mutation,summarize,count=length(mutation))

sup_4B<-ggplot(minor_counts,aes(x=count))+geom_histogram(color="white",binwidth=1)+xlab("Samples with iSNV")+ylab("Number of iSNV")
sup_4B+ggtitle("2005-2006")



pdf("./figures/sup_4B.pdf")
sup_4B
dev.off()



qual.chr.counts<-ddply(minor.05.06,~Intervention+chr+Id,summarize,iSNV=length(mutation))


columns<-unique(qual.chr.counts$Intervention)
qual.chr.counts$n=NA
for(i in 1:length(columns)){
x<-length(unique(qual.chr.counts$Id[which(qual.chr.counts$Intervention==columns[i])]))
qual.chr.counts$n[which(qual.chr.counts$Intervention==columns[i])]=x
}

qual.chr.counts<-mutate(qual.chr.counts,Intervention=paste0(Intervention," ","(n=",n,")"))


qual.chr.sum<-ddply(qual.chr.counts,~Intervention+chr,summarize,mean_iSNV=mean(iSNV),range=IQR(iSNV))

qual.chr.tab<-mutate(qual.chr.sum,iSNV=paste0(round(mean_iSNV,2)," \u00b1 ","(",range,")"))
qual.chr.table<-dcast(qual.chr.tab,chr~Intervention,value.var="iSNV")
knitr::kable(qual.chr.table)

qual.chr.counts<-ddply(minor.05.06,~HAI.geo+chr+Id,summarize,iSNV=length(mutation))

qual.chr.counts<-mutate(qual.chr.counts,HAI.geo=ifelse(HAI.geo,yes = "Above geometric mean",no= "Below geometric mean"))

columns<-unique(qual.chr.counts$HAI.geo)
qual.chr.counts$n=NA
for(i in 1:length(columns)){
x<-length(unique(qual.chr.counts$Id[which(qual.chr.counts$HAI.geo==columns[i])]))
qual.chr.counts$n[which(qual.chr.counts$HAI.geo==columns[i])]=x
}

qual.chr.counts<-mutate(qual.chr.counts,HAI.geo=paste0(HAI.geo," ","(n=",n,")"))


qual.chr.sum<-ddply(qual.chr.counts,~HAI.geo+chr,summarize,mean_iSNV=mean(iSNV),range=IQR(iSNV))

qual.chr.tab<-mutate(qual.chr.sum,iSNV=paste0(round(mean_iSNV,2)," \u00b1 ","(",range,")"))
qual.chr.table<-dcast(qual.chr.tab,chr~HAI.geo,value.var="iSNV")
knitr::kable(qual.chr.table)

```


# Appendix

```{r}
count_muts_samples<-function(data.df){
  data.df<-subset(data.df,freq.var<=0.5)
  out<-ddply(data.df,~Id,summarize,muts=length(mutation),HAI.WI=HAI.WI.30.post.vax[1],NAI.WI=NAI.WI.30.post.vax[1],Copy_num=Copy_num[1],dpi=Day.of.Infection.sample.collected[1])
}

lines.07<-count_muts_samples(minor.07.08)


ggplot(lines.07,aes(y=muts,x=Copy_num))+geom_point()
ggplot(lines.07,aes(y=muts,x=HAI.WI))+geom_point()
ggplot(lines.07,aes(y=muts,x=NAI.WI))+geom_point()
ggplot(lines.07,aes(y=muts,x=dpi))+geom_point()


```


```{r,by_day}
isnv_day<-function(df){
  df<-subset(df,freq.var<=0.5)
  ddply(df,~Id+Day.of.Infection.sample.collected+Intervention,summarize,iSNV=length(mutation))
}

#all.isnv.day<-isnv_day(var.2007.8.df)
#all.isnv.day$data="All"
qual.isnv.day<-isnv_day(minor.07.08)
#qual.isnv.day$data="High Quality"

#all.data.isnv.day<-rbind(all.isnv.day,qual.isnv.day)
qual.isnv.day<-rbind(qual.isnv.day,data.frame(Id=rep("extra",3),Day.of.Infection.sample.collected=c(5,6,6),Intervention=c("IIV","IIV","LAIV"),iSNV=rep(NA,3))) # this keeps everything in order. We don't had data for these days and interventions


iSNV.day<-ggplot(data=qual.isnv.day,aes(y=iSNV,x=as.factor(paste(Day.of.Infection.sample.collected,Intervention,sep="_"))))+geom_dotplot(aes(fill=Intervention),stackdir = "center",binaxis = 'y',position='dodge',binwidth = 1,dotsize = 0.7)+ylab("Number of SNV")+ggtitle("iSNV by day")+xlab("Day of Infection")+stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,geom = "crossbar", width = 0.5)+scale_y_continuous(limits=c(0,18),breaks=seq(0,18,2))+scale_x_discrete(labels=c("","2","","","3","","","4","","","5","","","6",""))

# pdf("~/Box Sync/FluVax Paper/Graphs/iSNV.day.pdf",width=9,height=4)
# iSNV.day
# dev.off()

```



It looks like the iSNV count peaks around day 3 or 4. But I don't know if that's statistically significant or robust.

