---
title: "Fluvacs Processing"
author: "JT McCrone"
date: "July 7, 2016"
output:
  pdf_document: default
  html_document: default
---

This document will run the first analysis on the variant calls. It will take in the csv files from the variant calling pipeline and filter out low quality calls as well as handle any samples done in duplicate. This script will then output 3 csv files. One that contains all minor quality variants. This will be used to make the figures. The second will just include quality calls from the HA segement. This will be used to identify putative antigenic variants. These go in the results directory as they are probably interesting for others as tables. The third is an improved metadata csv which includes information regarding the geometric mean of antibody titer and the day of collection in year.decimal format.
```{r,include=F}
require(knitr)
require(plyr)
require(ggplot2)
require(reshape2)
require(ggplot2)
require(ggdendro)
require(grid)
opts_chunk$set(fig.align="center",warning=FALSE,tidy=T,cache = F,echo=T)
theme_set(new = theme_classic()+ theme(
axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))) # to make nice plots
source("../scripts/useful_functions.R") # useful functions adapted largley from HIVE work
```

Read in the meta data file
```{r}
meta.all<-read.csv("../data/raw/all.meta.csv",stringsAsFactors = F)
```

Here I read in the variants calls, filter the calls for quality, apply the duplicate analysis on the appropriate sample, and trim for coding regions.

```{r, variant_calls}
## Read in and process samples
var.2007.8<-read_rbind(c("../data/processed/Run_1293/Variants/all.sum.csv","../data/processed/Run_1304/Variants/all.sum.csv","../data/processed/2007-2008/Variants/all.sum.csv"),c("Run_1293","Run_1304","2007-2008"))


var.2007.8.df<-processing(data.df = var.2007.8,meta.df = meta.all,pval = 0.01,phred = 35,mapq = 30,read_cut = c(32,94),gc=1e3)

ddply(subset(var.2007.8.df, Id=="34" & freq.var<0.5 & freq.var>0.01),~run,summarize, length(mutation))

minor.07.08<-quality(var.2007.8.df,c(0.01,0.5))

# Processing again with 34 only once
var.2007.8.df.cut<-subset(var.2007.8.df,!(Id==34 & run=="2007-2008"))

minor.07.08<-quality(var.2007.8.df.cut,c(0.01,0.5))


bris.bed<-read.csv("../data/processed/bis_difference.csv",stringsAsFactors = F,comment.char = "#") 

# Adjust the coding positions of each variant
coding.adjust.bris<-function(x){
  chr<-unique(x$chr)
  start<-bris.bed$off.5[match(x$chr,bris.bed$chr)]

  mutate(x,coding.pos=pos-start)
  }


minor.07.08<-ddply(minor.07.08,~chr,coding.adjust.bris)

## Other seasons
## Read in and process samples
other.seasons<-read_rbind(c("../data/processed/Run_1412/Variants/all.sum.csv","../data/processed/2005-2006/Variants/all.sum.csv","../data/processed/2004-2005/Variants/all.sum.csv"),c("Run_1412","2005-2006","2004-2005"))


var.other.seasons.df<-processing(data.df = other.seasons,meta.df = meta.all,pval = 0.01,phred = 35,mapq = 30,read_cut = c(32,94),gc=1e3)


ddply(subset(var.other.seasons.df, Id %in% c("529","530") & freq.var<0.5 & freq.var>0.01),~run+Id,summarize, length(mutation))

minor.other.seasons<-quality(var.other.seasons.df,c(0.01,0.5))


# Processing again with 529 and 530 only once
var.other.seasons.df.cut<-subset(var.other.seasons.df,!(Id==530 & run=="2005-2006"))

var.other.seasons.df.cut<-subset(var.other.seasons.df.cut,!(Id==529 & run=="2005-2006"))

minor.other.seasons<-quality(var.other.seasons.df.cut,c(0.01,0.5))


bris.bed<-read.csv("../data/processed/bis_difference.csv",stringsAsFactors = F,comment.char = "#") 


# Filter to coding region.



cal.bed<-read.csv("../data/processed/CalH3N2_difference.csv")
coding.adjust.cal<-function(x){
  chr<-unique(x$chr)
  start<-cal.bed$off.5[match(x$chr,cal.bed$chr)]

  mutate(x,coding.pos=pos-start)
  }

minor.other.seasons<-ddply(minor.other.seasons,~chr,coding.adjust.cal)

minor.04.05<-subset(minor.other.seasons,season=="04-05")
minor.05.06<-subset(minor.other.seasons,season=="05-06")



```

Initially we partitioned the data relative to the geometric mean of the antibody titer for a given season. In the final anlysis we used a titer of 40 as the cutoff. These cut offs were identical except for in 2005-2006 in which the mean was 30. This affected 2 samples.
```{r}
gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){ # from http://stackoverflow.com/questions/2602583/geometric-mean-is-there-a-built-in
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}





### For consensus seqeunce meta data

HAI.gm<-ddply(meta.all,~season,summarise,gm_mean=gm_mean(HAI.post.vax))
NAI.gm<-ddply(meta.all,~season,summarise,gm_mean=gm_mean(NAI.post.vax))

HAI.cut<-function(x){
  season<-unique(x$season)
  gm<-HAI.gm$gm_mean[which(HAI.gm$season==season)]
  print(gm)
  mutate(x,HAI.geo=HAI.post.vax>gm,HAI.cut=HAI.post.vax>=40)
  }

NAI.cut<-function(x){
  season<-unique(x$season)
  gm<-NAI.gm$gm_mean[which(NAI.gm$season==season)]
  print(gm)
  mutate(x,NAI.geo=NAI.post.vax>gm,NAI.cut=NAI.post.vax>=40)
  }
meta.df<-ddply(meta.all,~season,HAI.cut)
meta.df<-ddply(meta.df,~season,NAI.cut)


minor.04.05<-ddply(minor.04.05,~season,HAI.cut)
minor.04.05<-ddply(minor.04.05,~season,NAI.cut)

minor.05.06<-ddply(minor.05.06,~season,HAI.cut)
minor.05.06<-ddply(minor.05.06,~season,NAI.cut)

minor.07.08<-ddply(minor.07.08,~season,HAI.cut)
minor.07.08<-ddply(minor.07.08,~season,NAI.cut)


#meta.df<-mutate(meta.all,HAI.geo=HAI.WI.30.post.vax>gm_mean(HAI.WI.30.post.vax),NAI.geo=NAI.WI.30.post.vax>gm_mean(NAI.WI.30.post.vax))
meta.df$collection_date<-as.Date(meta.df$collection_date,format = "%d-%b-%y")
require(lubridate)
meta.df$collection_date<-decimal_date(meta.df$collection_date)


```

writing the outputs

```{r,eval=T}
## Getting data for HA
#ha<-subset(qual.2007.8,chr=='HA')
ha.07.08<-subset(minor.07.08,chr=='HA')
ha.04.05<-subset(minor.04.05,chr=='HA')
ha.05.06<-subset(minor.05.06,chr=='HA')



# dim(ha)
# ha<-ha[order(ha$pos),]
# unique(ha$pos)
write.csv(x = ha.07.08,file="../results/2007-2008.HA.csv",row.names = F)
write.csv(x = ha.04.05,file="../results/2004-2005.HA.csv",row.names = F)
write.csv(x = ha.05.06,file="../results/2005-2006.HA.csv",row.names = F)


write.csv(x=minor.07.08,file="../results/2007-2008.wg.csv",row.names=F)
write.csv(x=minor.04.05,file="../results/2004-2005.wg.csv",row.names=F)
write.csv(x=minor.05.06,file="../results/2005-2006.wg.csv",row.names=F)
write.csv(meta.df,"../data/raw/meta.all.HAgm.csv") 
```


